---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Loading the data file
```{r, echo = TRUE}
RR_week2 <- read.csv("activity.csv", na.strings=c("NA"))
```

### Preprocessing the data file
```{r, echo = TRUE}
# check if variables are in correct format
str(RR_week2)

# make steps numeric
steps <- as.numeric(RR_week2$steps, na.rm = TRUE)
RR_week2$steps <- steps 
```

## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day
```{r, echo = TRUE}
# load dplyr
library(dplyr)

# calculate the total number of steps per day
total_per_day <- RR_week2 %>%
  group_by(date) %>%
  summarise(steps_per_day = sum(steps))
```

### Create a histogramm of the total number of steps taken per day

```{r, echo=TRUE}
# plot a histogram with the total number of steps per day
ggplot(total_per_day, aes(x=date, y=steps_per_day)) + geom_histogram(stat="identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Calculate and report the mean and median of the total number of steps taken per day 

```{r, echo=TRUE}
# calculate the mean number of total steps per day
#mean_per_day <- RR_week2 %>%
#  group_by(date) %>%
#  summarise(steps_per_day = mean(steps))
#mean_per_day

#median_per_day <- RR_week2 %>%
#  group_by(date) %>%
#  summarise(steps_per_day = median(steps))
#median_per_day

mean_total_steps_across_all_days <- mean(total_per_day$steps, na.rm=TRUE)
mean_total_steps_across_all_days

median_total_steps_across_all_days <- median(total_per_day$steps, na.rm=TRUE)
median_total_steps_across_all_days
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
